---
title: "Lab 06 - Functions and iterations"
author: "Kaitlin Fong"
date: "`r Sys.Date()`"
output: html_document
---


### Packages and Data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(ggplot2)
```

```{r diamon_rename}
diamonds <- diamonds %>% rename(
    depth_pct = depth,
    length_mm = x,
    width_mm = y,
    depth_mm = z,
    table_pct = table
  )
```


## Exercise 1

```{r}
# Do some preliminary investigations & cleaning
ggplot(data = diamonds,
       mapping = aes(x = width_mm,
                     y = depth_mm)) +
  geom_point()

diamonds <- diamonds %>% filter(
  length_mm != 0, length_mm <= 20,
  width_mm != 0, width_mm <= 20,
  depth_mm != 0, depth_mm <= 20
)
```


## Exercise 2

```{r}

diamonds <- diamonds %>%
  mutate(table_mm = table_pct * width_mm / 100)

# Overall table_mm arithmetic mean?

diamonds %>%
  summarise(avg_table = mean(table_mm, na.rm = T))




# table_mm arithmetic mean within each clarity category

diamonds %>%
  group_by(clarity) %>%
  summarise(avg_table = mean(table_mm, na.rm = T))




```

_The table mean decreases as the clarity of the diamonds increase (from included to flawless)_

## Exercise 3

_Do derivation on paper, unless you know how to write LaTeX code_

## Exercise 4

```{r}
# Write gmean() command
gmean <- function(x){
  xbar_g <- exp(mean(log(x)))
  return(xbar_g)
}
```

## Exercise 5

```{r, error = TRUE}

test_data <- list(
  test1 = c(2.1, 3.8, 4.2),
  test2 = c(1, 10, 100, 1000),
  test3 = c(0, 1, 4),
  test4 = c(0.38,  0.94, -1.56),
  test5 = c(TRUE, TRUE),
  test6 = c("6", "7", "8")
)


# Create for loop to test gmean() on the above examples
for(i in seq_along(test_data)){
  print(gmean(test_data[[i]]))
}
```


## Exercise 6

```{r}
# Copy gmean() from Ex4 and edit it to check the input
gmean <- function(x){
  
  if(!is.numeric(x)){
    warning("Vector 'x' is not numeric")
    return(NaN)
  }
    
  if(any(x <= 0)){
    warning("Vector 'x' must have strictly positive numbers.")
    return(NaN)
  }
  
  xbar_g <- exp(mean(log(x)))
  return(xbar_g)
}





# Copy your code from Ex 5 to test your new command

for(i in seq_along(test_data)){
  print(gmean(test_data[[i]]))
}



```


## Exercise 7

```{r}
# Overall table_mm geometic mean?

diamonds %>% 
  summarise(
    avg_table = gmean(table_mm)
  )


# table_mm geometic mean within each clarity category

diamonds %>% 
  group_by(clarity) %>%
  summarise(
    avg_table = gmean(table_mm)
  )

```

_Diamonds with a better clarity have a smaller geometric mean. Also, the geometric means are typically smaller than the arithmetic mean for this data._


## Exercise 8 (optional)

```{r}
# Create gmean2() that computes the geometric mean 
#   that uses the original definition

gmean2 <- function(x){
  
  if(!is.numeric(x)){
    warning("Vector 'x' is not numeric")
    return(NaN)
  }
    
  if(any(x <= 0)){
    warning("Vector 'x' must have strictly positive numbers.")
    return(NaN)
  }
  
  xbar_g <- prod(x)^(1/length(x))
  return(xbar_g)
}

gmean(test_data$test1)

```


## Exercise 9 (optional)

```{r}
# Create a function that computes the harmonic mean

hmean <- function(x){
  
  if(!is.numeric(x)){
    warning("Vector 'x' is not numeric")
    return(NaN)
  }
    
  if(any(x <= 0)){
    warning("Vector 'x' must have strictly positive numbers.")
    return(NaN)
  }
  
  xbar_h <- 1 / mean(1 / x)
  return(xbar_h)
}

diamonds %>% 
  summarise(
    avg_table = hmean(table_mm)
  )

```
